\section{Mathematical Foundations}\label{sec:math_foundations}

\subsection{Problem Formulation}\label{subsec:problem_formulation}
Assuming we are given a set of keypoints corresponding to the same physical points in both images, the localization task requires establishing a mathematical mapping between the UAV's local perspective and the global reference map. We define the following variables:

\begin{itemize}
    \item $\mathbf{p}_i^{uav} = [x_i, y_i, 1]^T$: A homogeneous keypoint in the UAV-image coordinate system. ($1$ as the third coordinate is used since at the current moment we are considering only similarity and affine transformations)
    \item $x_i, y_i$: The pixel coordinates of a feature in the UAV camera feed.
    \item $\mathbf{p}_i^{map} = [X_i, Y_i, 1]^T$: The corresponding homogeneous keypoint in the map-image coordinate system.
    \item $X_i, Y_i$: The pixel coordinates of the matching feature on the satellite map.
    \item $n$: The total number of matched keypoint pairs.
    \item $T$: The transformation matrix mapping the UAV coordinates to the map coordinates.
\end{itemize}

Given the set of $n$ matched pairs, the fundamental goal is to find the matrix $T$ such that:
\begin{equation}
\mathbf{p}_i^{map} \approx T\mathbf{p}_i^{uav}.
\end{equation}

The problem is posed as an optimization task. The objective is to estimate the optimal transformation matrix $\hat{T}$ that minimizes the reprojection error across all $N$ correspondences:
\begin{equation}
\hat{T} = \arg\min_T \sum_{i=1}^{n} \left\|\mathbf{p}_i^{map} - T\mathbf{p}_i^{uav}\right\|^2.
\end{equation}



\subsection{Scale-Invariant Feature Detection and Matching}
To establish correspondences between the UAV imagery and the reference map, we utilize the Scale-Invariant Feature Transform (SIFT) algorithm, which represents keypoints as 128-dimensional continuous feature vectors. The similarity between a query descriptor $v$ and a train descriptor $u$ is determined using the Euclidean distance ($L_2$ norm) in the feature space:
$$d(v, u) = \sqrt{\sum_{k=1}^{128} (v_k - u_k)^2}$$
Due to the highly repetitive nature of aerial landscapes, direct nearest-neighbor matching is prone to false positives. To mathematically mitigate this, correspondences are filtered using Lowe's ratio test. For any given descriptor, let $d_1$ be the distance to its closest neighbor and $d_2$ be the distance to the second closest neighbor. A match is considered reliable and unique only if it satisfies the following inequality:
$$\frac{d_1}{d_2} < \tau$$
where $\tau$ is a predefined ratio threshold.

\subsection{Robust Estimation via RANSAC}

Given a set of noisy correspondences between UAV image points and map points,
\[
\mathcal{D} = \{(x_i, y_i) \leftrightarrow (X_i, Y_i)\}_{i=1}^{N},
\]
the goal is to estimate transformation parameters $\theta$ (affine model)
that are consistent with the largest subset of correct matches (inliers),
while rejecting incorrect correspondences (outliers).

\textbf{Reprojection Error Definition.}

Each correspondence induces a reprojection error under model $\theta$:
\[
d_i(\theta) = \left\| 
\begin{bmatrix} X_i \\ Y_i \end{bmatrix}
-
\hat{T}_\theta
\begin{bmatrix} x_i \\ y_i \\ 1 \end{bmatrix}
\right\|_2,
\]
where $\hat{T}_\theta$ is the affine transformation matrix.

A point is considered an inlier if:
\[
d_i(\theta) < \epsilon,
\]
where $\epsilon$ is a predefined threshold.

Thus, the objective is:
\[
\theta^* = \arg\max_{\theta} \sum_{i=1}^{N} \mathbf{1}\big(d_i(\theta) < \epsilon\big),
\]
i.e., maximize the number of inliers.

\textbf{RANSAC Iterative Procedure.}

At each iteration, a minimal subset of $s$ correspondences is sampled:
\[
\mathcal{S} \subset \mathcal{D}, \quad |\mathcal{S}| = s,
\]
where for the affine model $s = 3$.

Using this subset, a candidate model $\theta$ is estimated by solving
the linear system defined by the affine transformation.

The model is then evaluated on the full dataset by computing
the reprojection errors $\{d_i(\theta)\}_{i=1}^{N}$ and constructing
the inlier set:
\[
\mathcal{I}(\theta) = \{ i \mid d_i(\theta) < \epsilon \}.
\]

The model with the largest consensus set is selected:
\[
\theta^* = \arg\max_{\theta} |\mathcal{I}(\theta)|.
\]

\textbf{Adaptive Number of Iterations.}

The number of RANSAC iterations is derived probabilistically.
Let $w$ denote the probability that a randomly selected correspondence is an inlier:
\[
w = \frac{|\mathcal{I}|}{N}.
\]

The probability that all $s$ points in a randomly sampled subset are inliers is:
\[
P_{\text{all-inliers}} = w^s.
\]

Thus, the probability that a sampled subset contains at least one outlier is:
\[
1 - w^s.
\]

If we perform $k$ independent iterations, the probability that all sampled subsets
contain at least one outlier is:
\[
(1 - w^s)^k.
\]

Therefore, the probability that at least one of the $k$ samples consists entirely
of inliers is:
\[
1 - (1 - w^s)^k.
\]

To guarantee a desired confidence level $p$, we require:
\[
1 - (1 - w^s)^k \geq p.
\]

Rearranging the inequality:
\[
(1 - w^s)^k \leq 1 - p.
\]

Taking logarithms on both sides:
\[
k \log(1 - w^s) \leq \log(1 - p).
\]

Finally, solving for $k$:
\[
k = \frac{\log(1 - p)}{\log(1 - w^s)}.
\]

\textbf{Final Model Estimation.}

After identifying the best inlier set $\mathcal{I}^*$, the final model
parameters are recomputed using all inliers by solving the
overdetermined system:
\[
A v = b,
\]
using the least-squares solution:
\[
v = (A^T A)^{-1} A^T b.
\]



\subsection{Geometric Transformation Models}\label{subsec:transformation_models}
Geometric transformations define the mathematical mapping between UAV imagery and the reference map. We consider three linear models:

\begin{itemize}
    \item \textbf{Similarity Transformation:} A 4-degree-of-freedom model accounting for translation, rotation, and uniform scaling.
    \item \textbf{Affine Transformation:} A 6-degree-of-freedom model extending similarity by adding shear and non-uniform scaling.
    \item \textbf{Projective Transformation:} An 8-degree-of-freedom model representing the true 3D to 2D projection.
\end{itemize}

The Similarity model requires perfectly level flight and uniform zooming, failing with axial stretching or tilt. The Affine model approximates slight perspective shifts via shear, but only the Projective model handles significant pitch and roll by modeling converging parallel lines.

At the current stage, we estimate transformations using the Similarity and Affine models.

The Similarity model is written as:
\begin{equation}
\begin{bmatrix} X \\ Y \\ 1 \end{bmatrix} =
\begin{bmatrix}
a & -b & t_x \\
b &  a & t_y \\
0 &  0 & 1
\end{bmatrix}
\begin{bmatrix} x \\ y \\ 1 \end{bmatrix},
\end{equation}
where $a = s\cos\theta$ and $b = s\sin\theta$.

For one correspondence $(x_i, y_i) \leftrightarrow (X_i, Y_i)$, this yields the linear equations
\begin{equation}
X_i = ax_i - by_i + t_x, \quad Y_i = bx_i + ay_i + t_y,
\end{equation}
which are stacked over all inliers into an overdetermined system:
\begin{equation}
\begin{bmatrix}
x_1 & -y_1 & 1 & 0 \\
y_1 &  x_1 & 0 & 1 \\
\vdots & \vdots & \vdots & \vdots \\
x_n & -y_n & 1 & 0 \\
y_n &  x_n & 0 & 1
\end{bmatrix}
\begin{bmatrix} a \\ b \\ t_x \\ t_y \end{bmatrix}
=
\begin{bmatrix} X_1 \\ Y_1 \\ \vdots \\ X_n \\ Y_n \end{bmatrix}.
\end{equation}

The Affine model uses the same construction principle (stacking two linear equations per correspondence), but with six unknowns and matrix form:
\begin{equation}
\begin{bmatrix} X \\ Y \\ 1 \end{bmatrix} =
\begin{bmatrix} a_1 & a_2 & t_x \\ a_3 & a_4 & t_y \\ 0 & 0 & 1 \end{bmatrix}
\begin{bmatrix} x \\ y \\ 1 \end{bmatrix}
\end{equation}

\subsection{Least-Squares Parameter Estimation}\label{subsec:least_squares}
Both Similarity and Affine models produce overdetermined linear systems of the form $A\mathbf{v}=\mathbf{b}$ with $2N$ equations. Because $\mathbf{b}$ rarely lies exactly within the column space of $A$, denoted $C(A)$, an exact solution is unlikely. Instead, the optimal estimate $\hat{\mathbf{v}}$ minimizes the residual error $\mathbf{r} = \mathbf{b} - A\mathbf{v}$. Geometrically, this requires finding the vector $A\hat{\mathbf{v}}$ in $C(A)$ that is closest to $\mathbf{b}$, which is its orthogonal projection.

For $A\hat{\mathbf{v}}$ to be the precise orthogonal projection, the residual vector $\mathbf{r}$ must be strictly orthogonal to the entire column space of $A$. Consequently, the dot product of the residual with every column of $A$ must equal zero.
\begin{equation}
A^T \mathbf{r} = \mathbf{0}.
\end{equation}

Substituting the definition of the residual, $\mathbf{r} = \mathbf{b} - A\hat{\mathbf{v}}$, into this orthogonality condition yields:
\begin{equation}
A^T (\mathbf{b} - A\hat{\mathbf{v}}) = \mathbf{0}.
\end{equation}

Distributing $A^T$ across the terms gives:
\begin{equation}
A^T \mathbf{b} - A^T A \hat{\mathbf{v}} = \mathbf{0}.
\end{equation}

Rearranging this equality results in the fundamental normal equations:
\begin{equation}
A^T A \hat{\mathbf{v}} = A^T \mathbf{b}.
\end{equation}

Assuming the columns of $A$ are linearly independent, the square symmetric matrix $A^T A$ is invertible. Geometrically, this independence is guaranteed as long as there are at least three non-collinear matched points, a condition easily satisfied by the vast number of dispersed keypoints typically extracted from the images. Multiplying both sides by its inverse provides the final closed-form estimator:
\begin{equation}
\hat{\mathbf{v}} = (A^T A)^{-1}A^T \mathbf{b}.
\end{equation}

This geometrical formulation is entirely model-agnostic: only the structure of $A$ and the length of $\mathbf{v}$ change between the Similarity and Affine models.